State of the Practice for Medical Imaging Software

### 2022-03-26 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Updated AHP figures
Edits to paper

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
	How do MI projects compare to research software in general with respect to
	the use of tools (Section~\ref{Sec_CompareTools}) for:
	\begin{enumerate} 
		\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
		\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
	\end{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
	How do MI projects compare to research software in general with respect to
	principles, processes and methodologies used?
	(Section~\ref{Sec_CompareMethodologies})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on MI software projects?
	(Section~\ref{ch_interview})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from MI compare to the pain points
	for research software in general? (Section~\ref{ch_interview})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
	developers what specific best practices are taken to address the pain points
	and software quality concerns? (Section~\ref{ch_interview})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practice could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}).
	(Section~\ref{ch_recommendations})

@@

@@

\section{Interviews with Developers} \label{ch_interview}

@@

\begin{description}
\item[Q20.] Do you have any concern that your computational results won’t be
reproducible in the future? Have you taken any steps to ensure reproducibility?
\end{description}

@@

\section{Answers to Research Questions} \label{ch_answers}

This section answers the research questions from
Section~\ref{sec_motivation}.  We will go through each of the questions
in turn.

\begin{description}
\item[RQ2.] What tools are used in the development of current software packages?
\end{description}

\begin{description}
\item[RQ3.] What principles, processes, and methodologies are used in the
development of current software packages?
\end{description}

We answer this question by measuring the qualities and interviewing the
developers. We previously summarized principles, processes, and methodologies in
software testing (Section~\ref{sec_interview_pain_points} and
Section~\ref{sec_interview_software_qualities}), documentation
(Section~\ref{sec_interview_documents}, Table~\ref{tab_maintainability_docs},
Table~\ref{tab_Visibility/Transparency_docs}, and Section~\ref{ch_answers}),
contribution management (Section~\ref{sec_contribution_pm}), and project
management (Section~\ref{sec_contribution_pm}).

We identified the use of unit testing in less than half of the 29 projects. On
the other hand, the interviewees believed that testing (including usability
tests with users) was the top solution to improve correctness, usability, and
reproducibility. One pain point in the development process is the lack of access
to real-world datasets for testing. The developers' strategies to address this
are summarized in Section~\ref{ch_interview} \ad{to be fixed}. One threat to correctness is:
with huge datasets for testing, the tests are expensive and time-consuming.
Three interviewees endorsed self tests / automated tests, which may save time
for testing.

\begin{description}
\item[RQ4.] What are the pain points for developers working on research software
projects? What aspects of the existing processes, methodologies, and tools do
they consider as potentially needing improvement? What changes to processes,
methodologies, and tools can improve software development and software quality?
\end{description}

We answer this question by interviewing the developers. The answers to this
question, including the pain points and the solutions proposed by the
developers, are in Section~\ref{sec_interview_pain_points}.

\begin{description}
\item[RQ5.] What is the current status of the following
software qualities for the projects? What actions have the developers taken to
address them?
\end{description}

The current status of the software qualities was assessed through filling in the
measurement template \citep{Dong2021-Data}.  A summary of the comparison between
all projects can be found in Section~\ref{ch_results}.  We explored some of the
qualities more deeply, especially with respect to techniques addressing quality
concerns, through the interview questions.  Specifically, the qualities of
correctness \& verifiability, usability, maintainability, understandability and
reproducibility are explored in Section~\ref{ch_interview}.

\begin{description}
\item[RQ6.] How does software designated as high quality by this methodology
compare with top-rated software by the community?
\end{description}

@@

\begin{itemize}
\item For practical considerations the time spent measuring each package had to
be limited.  The time limit may have caused an assessment to have missed
something something relevant.
\item Our ranking is partly based on surface (shallow) measurement, which may
not fully reveal the underlying qualities.
\item The questions in the measurement template may not actually measure the
qualities they are associated with.  For instance, we have assumed that
maintainability is improved if a high percentage of identified issues are
closed, but it is possible for a project with a wealth of ideas to have many
open issues, and still be maintainable.
\item With the exception of the interview data for 8 projects, we collected all
of the information for each project from the artifacts available on the
Internet. In some cases this source of data may mean we did not find evidence of
something, like unit testing, not because the project didn't do it, but because
no artifacts of this activity remained in the publicly available repository.
\item As mentioned in Section~\ref{ch_interview}, one interviewee was too busy
to participate in a full interview, so they provided a version of written answers
to us. Since we did not have the chance to explain our questions or ask them
follow-up questions, there is a possibility of misinterpretation of the
questions or answers.
\item As mentioned in Section~\ref{sec_result_installability}, we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We used
a deployed online version for \textit{dwv}, a VM version for \textit{GATE}, but
no alternative for \textit{DICOM Viewer}. We might underestimate their rank due
to an uncommon technical issue.
\end{itemize}

@@

This section presents our recommendations on MI software development. Although
our focus is on MI software, unless noted otherwise, our recommendations apply
to any scientific computing software.
Section~\ref{sec_recommendations_qualities} discusses the actions that can
potentially improve the ten software qualities.
Sections~\ref{sec_recommendations_limited_resources},
\ref{sec_recommendations_tech_stack}, and
\ref{sec_recommendations_testing_dataset} are based on the primary pain points
(from Section~\ref{sec_interview_pain_points}) collected from developers in the
MI domain.

@@

Based on our quality measurements in Section~\ref{ch_results} and discussions
with the developers in Section~\ref{ch_interview}, we collected key points for
improving software qualities. These points should be considered for new and
existing projects, but we are not saying that every project should do everything
on our list.  All projects will have finite resources, so strategies will have
to be selected that provide the greatest return on investment.  Our specific
recommendations for each quality are as follows:

@@

The limitation of resources has many faces, with the key manifestations being
lack of fundings, time, and developers. We summarized our discussion with the MI
software developers in Section~\ref{sec_interview_pain_points}.  Many of our
recommendations involve investing more effort into processes, tools and methods,
so that the limited resources can be used more effectively.  Our recommendations
are as follows:

@@

A tech stack refers to a set of technologies used by a team to build software
and manage the project. Section~\ref{sec_interview_pain_points} lists the
advantages and disadvantages between native and web applications. In this
section, we give further suggestions on the choice of a tech stack to improve
the four priority factors identified by developers: compatibility,
maintainability, performance and security.  The suggestions are intended to
provide ideas and avenues for exploration; not all of the suggestions will be
the right fit for all projects and all teams.

@@

As described in Section~\ref{sec_interview_pain_points}, it is difficult for
software development teams to access real-world medical imaging datasets. This
problem restricts their capability and flexibility for testing. We provide some
suggestions as follows:

@@

We analyzed the state of the practice for the MI domain with the goal of
understanding current practice, answering our six research questions
(Section~\ref{sec_motivation}) and providing recommendations for current
and future projects.  Our methods in Section~\ref{ch_methods} form a general
process to evaluate domain-specific software, that we apply to the specific
domain of MI software. We identified 48 MI software candidates, then, with the
help of the Domain Expert selected 29 of them to our final list.
Section~\ref{ch_results} lists our measurements to nine software qualities for
the 29 projects, and Section~\ref{ch_interview} contains our interviews with
eight of the 29 teams, discussing their development process and five software
qualities.  We answered our six research questions in Section~\ref{ch_answers}.
In addition, Section~\ref{ch_recommendations} presents our recommendations on SC
software development.

@@

The interview results in Section~\ref{ch_interview} show some merits, drawbacks,
and pain points within the development process. The three primary categories of
pain points are:
\begin{itemize}
\item the lack of fundings and time;
\item the difficulty to balance between four factors: cross-platform
compatibility, convenience to development \& maintenance, performance, and
security;
\item the lack of access to real-world datasets for testing.
\end{itemize}
We summarized the solutions from the developers to address these problems,
including developing a web-based approach with backend servers and maintaining
better documentation. We also collected the status of documentation.  We found
that for all 8 interviewed teams that documentation is felt to be vital to a
project, with the most popular form of documentation being forum discussions and
videos.  With respect to project management almost all teams used GitHub and
pull requests to manage contributions.  Very few teams used a specific
development model.  It appears that the development process is more ad hoc than
planned for the majority of projects.

Our answers to the research questions (Section~\ref{ch_answers}) are based on
the above findings. We identified the existing artifacts, tools, principles,
processes, and methodologies in the 29 projects. By comparisons with the implied
popularity of existing projects we found: 1) four of the top five software
projects in our ranking were also among the top five ones receiving the most
GitHub stars per year (Table~\ref{tab_ranking_vs_GitHub}); 2) three of the top
four in our ranking were among the top four provided by the domain experts.

@@

>>>>

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
	How do MI projects compare to research software in general with respect to
	the use of tools (Section~\ref{Sec_CompareTools}) for:
	\begin{enumerate} 
		\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
		\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
	\end{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
	How do MI projects compare to research software in general with respect to
	principles, processes and methodologies used?
	(Section~\ref{Sec_CompareMethodologies})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on MI software projects?
	(Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from MI compare to the pain points
	for research software in general? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
	developers what specific best practices are taken to address the pain points
	and software quality concerns? (Section~\ref{Sec_AddressConcerns})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practice could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}).
	(Section~\ref{ch_recommendations})

@@

We answer this question by measuring the qualities and interviewing the
developers. We previously summarized principles, processes, and methodologies in
software testing (Section~\ref{Sec_CompareMethodologies}), documentation
(Section~\ref{sec_interview_documents}, Table~\ref{tab_maintainability_docs},
Table~\ref{tab_Visibility/Transparency_docs}, contribution management
(Section~\ref{sec_contribution_pm}), and project management
(Section~\ref{sec_contribution_pm}).

We identified the use of unit testing in less than half of the 29 projects. On
the other hand, the interviewees believed that testing (including usability
tests with users) was the top solution to improve correctness, usability, and
reproducibility. One pain point in the development process is the lack of access
to real-world datasets for testing. The developers' strategies to address this
are summarized in Section~\ref{Sec_AddressConcerns} \ad{to be fixed}. One threat to
correctness is: with huge datasets for testing, the tests are expensive and
time-consuming. Three interviewees endorsed self tests / automated tests, which
may save time for testing.

@@

@@

\begin{description}
\item[Q20.] Do you have any concern that your computational results won't be
reproducible in the future? Have you taken any steps to ensure reproducibility?
\end{description}

@@

\section{Lessons from MI Developers} \label{Sec_AddressConcerns}

This section summarizes the best practices from LBM developers that are taken to
address the pain points mentioned in Section~\ref{painpoints}.  The main source
of information is the qualitative data from developer interviews
(Section~\ref{sec_interview_methods}).  The practices summarized in this section
can potentially be emulated by the LBM software packages that do not currently
follow them.  Moreover, these practices may also provide examples that can be
followed by other research software domains.

@@

\begin{itemize}
\item For practical considerations the time spent measuring each package had to
be limited.  The time limit may have caused an assessment to have missed
something something relevant.
\item Our ranking is partly based on surface (shallow) measurement, which may
not fully reveal the underlying qualities.
\item The questions in the measurement template may not actually measure the
qualities they are associated with.  For instance, we have assumed that
maintainability is improved if a high percentage of identified issues are
closed, but it is possible for a project with a wealth of ideas to have many
open issues, and still be maintainable.
\item With the exception of the interview data for 8 projects, we collected all
of the information for each project from the artifacts available on the
Internet. In some cases this source of data may mean we did not find evidence of
something, like unit testing, not because the project didn't do it, but because
no artifacts of this activity remained in the publicly available repository.
\item As mentioned in Section~\ref{painpoints}, one interviewee was too busy
to participate in a full interview, so they provided a version of written answers
to us. Since we did not have the chance to explain our questions or ask them
follow-up questions, there is a possibility of misinterpretation of the
questions or answers.
\item As mentioned in Section~\ref{sec_result_installability}, we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We used
a deployed online version for \textit{dwv}, a VM version for \textit{GATE}, but
no alternative for \textit{DICOM Viewer}. We might underestimate their rank due
to an uncommon technical issue.
\end{itemize}

@@

This section presents our recommendations on MI software development. Although
our focus is on MI software, unless noted otherwise, our recommendations apply
to any scientific computing software.
Section~\ref{sec_recommendations_qualities} discusses the actions that can
potentially improve the ten software qualities.
Sections~\ref{sec_recommendations_limited_resources},
\ref{sec_recommendations_tech_stack}, and
\ref{sec_recommendations_testing_dataset} are based on the primary pain points
(from Section~\ref{painpoints}) collected from developers in the
MI domain.

@@

Based on our quality measurements in Section~\ref{ch_results} and discussions
with the developers in Section~\ref{painpoints}, we collected key points for
improving software qualities. These points should be considered for new and
existing projects, but we are not saying that every project should do everything
on our list.  All projects will have finite resources, so strategies will have
to be selected that provide the greatest return on investment.  Our specific
recommendations for each quality are as follows:

@@

The limitation of resources has many faces, with the key manifestations being
lack of fundings, time, and developers. We summarized our discussion with the MI
software developers in Section~\ref{painpoints}.  Many of our recommendations
involve investing more effort into processes, tools and methods, so that the
limited resources can be used more effectively.  Our recommendations are as
follows:

@@

A tech stack refers to a set of technologies used by a team to build software
and manage the project. Section~\ref{painpoints} lists the advantages and
disadvantages between native and web applications. In this section, we give
further suggestions on the choice of a tech stack to improve the four priority
factors identified by developers: compatibility, maintainability, performance
and security.  The suggestions are intended to provide ideas and avenues for
exploration; not all of the suggestions will be the right fit for all projects
and all teams.

@@

As described in Section~\ref{painpoints}, it is difficult for
software development teams to access real-world medical imaging datasets. This
problem restricts their capability and flexibility for testing. We provide some
suggestions as follows:

@@

We analyzed the state of the practice for the MI domain with the goal of
understanding current practice, answering our six research questions
(Section~\ref{sec_motivation}) and providing recommendations for current and
future projects.  Our methods in Section~\ref{ch_methods} form a general process
to evaluate domain-specific software, that we apply to the specific domain of MI
software. We identified 48 MI software candidates, then, with the help of the
Domain Expert selected 29 of them to our final list. Section~\ref{ch_results}
lists our measurements to nine software qualities for the 29 projects, and
Section~\ref{painpoints} contains our interviews with eight of the 29 teams,
discussing their development process and five software qualities.  We answered
our research questions. In addition, Section~\ref{ch_recommendations} presents
our recommendations on SC software development.

@@

The interview results in Section~\ref{painpoints} show some merits, drawbacks,
and pain points within the development process. The three primary categories of
pain points are:
\begin{itemize}
\item the lack of fundings and time;
\item the difficulty to balance between four factors: cross-platform
compatibility, convenience to development \& maintenance, performance, and
security;
\item the lack of access to real-world datasets for testing.
\end{itemize}
We summarized the solutions from the developers to address these problems,
including developing a web-based approach with backend servers and maintaining
better documentation. We also collected the status of documentation.  We found
that for all 8 interviewed teams that documentation is felt to be vital to a
project, with the most popular form of documentation being forum discussions and
videos.  With respect to project management almost all teams used GitHub and
pull requests to manage contributions.  Very few teams used a specific
development model.  It appears that the development process is more ad hoc than
planned for the majority of projects.

Our answers to the research questions are based on the above findings. We
identified the existing artifacts, tools, principles, processes, and
methodologies in the 29 projects. By comparisons with the implied popularity of
existing projects we found: 1) four of the top five software projects in our
ranking were also among the top five ones receiving the most GitHub stars per
year (Table~\ref{tab_ranking_vs_GitHub}); 2) three of the top four in our
ranking were among the top four provided by the domain experts.

@@

- add discussion of sensitivity analysis (borrow from LBM)
